\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-08-02 forwards-and-backwards.tex
.

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction},
For each value of $N$ and $\theta = \theta(N)$,
we define $(\eta^N_t)_{t \geq 0}$ to be the solution to the Martingale Problem,
with initial condition $\eta^N_0 \in \measures$,
such that for all smooth functions $f \in C^{\infty}_{0}(\mathbb{R}^d)$,
writing $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$
and $q_\theta(x, dy)$ as the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eq: Pre-limit Martingale Characterisation}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
\\ & \qquad \qquad \qquad \qquad {}
        \times \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N_t(f)] =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \langle \gamma(x, \eta^N_{s})
        \int f^2(y)r(y, \eta^N_{s}) q_\theta(x,dy) 
    , \eta^N_{s}(dx) \rangle \\
& \qquad \qquad {}
    + \langle \mu(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}

Note that
since individuals are produced at rate $N \gamma \eta$ but have mass $1/N$ each,
these factors of $N$ cancel in~\eqref{eq: Pre-limit Martingale Characterisation}.
Since we are taking a diffusion limit,
we will have use for the following notation:

@@

The lookdown processes have been carefully constructed so that
if we average over levels in the expression for the lookdown generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we get the generator for the population process
(visible in equation~\eqref{eq: Pre-limit Martingale Characterisation}
or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (and some boundedness conditions)
in Appendix \ref{sec: Markov Mapping Theorem Application},
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2018}
tells us that for any population process $(\eta^N_t)_{t \ge 0}$
(i.e., solution to the martingale problem of Definition~\ref{defn:mgale_construction})
% of equations~\eqref{eq: Pre-limit Martingale Characterisation} and \eqref{eqn:prelimit_martingale_variation}
there is a lookdown representation $(\lp^N_t)_{t \ge 0}$
(i.e., solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])$
and such that at each $t$, conditioned on $(\eta^N_s){s \le t}$, the levels of $\lp_t^N$
are independently and uniformly distributed on $[0, N]$.
The same property holds in the limit:
for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of Theorem~\ref{nonlocal_convergence}
there is a lookdown process $(\lp^N_t)_{t \ge 0}$
that is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
and such that at each $t$,
$\lp_t$ is conditionally Poisson with Cox measure $\eta_t(dx) \times du$,
conditioned on $(\eta_s)_{s \le t}$.
Furthermore, convergence of the lookdown processes would imply convergence of the population processes
\citep[Thm A.9 in][]{KR}.

@@

We prove tightness of $\eta^N$ as $\mathcal{D}_{\cmeasures}([0,\infty))$-valued processes,
where $\cmeasures$ is the space of finite measures on the one-point compactification of $\IR^d$.
To do this, we first prove that $\sup_{0 \le t \le T} \langle 1, \eta^N_t \rangle$ is tight
(a compact containment condition, Lemma~\ref{lem: Compact Containment}).
(The reason we use $\cmeasures$ instead of $\measures$
is so that the set of measures with total mass less than $K$ is compact.)
Next, in Lemma~ we show that the one-dimensional distributions,
$\langle f, \eta^N_t \rangle$ are tight
for each $f \in \mathcal{C}^2_0(\overline{\IR^d})$.
Since this is a dense subset of bounded continuous functions on $\cmeasures$,
together these results allow us to apply Theorem~3.9.1 of \citet{EK}.

@@

To show compact containment, we will need boundedness
of the total net reproduction rate.
This does not follow immediately from our assumption that $F$ is bounded above
because reproduction is not local.
\comment{say more about the difference between $F$ and the total net reproduction rate}
The following lemma will be useful.

\begin{lemma}
    \label{lem:q_bound}
    Under the assumptions of Definition~\ref{def:model_setup},
    for all $f \in \mathcal{C}^{2}_{0}(\IR^d \times [0,\infty))$,
    then there exists a $C_f < \infty$ such that
    for all $\theta > 0$ and all $\eta \in \measures$,
    \begin{equation*}
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
        \leq
        C_f \left( 1 + \smooth{\gamma} \eta(x)^2 \right)
    \end{equation*}
\end{lemma}

\begin{proof}[of Lemma \ref{lem:q_bound}]
    The key assumption here from Definition~\ref{def:model_setup}
    will be that $\epsilon_{r}^2 + \frac{2\lambda}{\theta} < \epsilon_{\gamma}^2$,
    where $\lambda = \sup_x \sup_{y : \|y\| = 1} y^T \covq(x) y$
    is the largest eigenvalue of $\covq(x)$ across all $x$.
    First note that 
    \begin{align} \label{eqn:q_bound_split}
    &
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(y)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(x, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy) 
        \right| .
    \end{align}

@@

    First we prove \eqref{eqn:first_moment_rho}.
    Recall that $\smooth{r} \eta(x) = \int p_{\epsilon^2_r}(x - w) \eta(dw)$,
    where $p_t$ is the density of a Gaussian with mean 0 and variance $t$,
    so that applying Fubini, \eqref{eqn:first_moment_rho} is
    \begin{align*}
        \left|
        \int_{\IR^d}
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            q_\theta(x, dy)
        \eta(dw)
        \right| .
    \end{align*}
    Write $p_{s, x}(\cdot)$ for the density of a Gaussian
    with mean $s \meanq(x)$ and covariance $\epsilon_r^2 I + s \covq(x)$,
    so that $\int p_{\epsilon_r^2}(y - w) q_\theta(x, dy) = p_{\epsilon_r^2, 1/\theta, x}(w-x)$.
    It therefore suffices to show that for all $x$ and $w \in \IR^d$,
    \begin{align*}
        \left|
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y-w) - p_{\epsilon_r^2}(x-w) )
            q_\theta(x, dy)
        \right|
        &=
        \theta \left|
            p_{1/\theta, x}(w-x) - p_{0, x}(w-x)
        \right|
        \le
        p_{\epsilon^2_\gamma}(w-x) .
    \end{align*}
    However,
    $\theta( p_{1/\theta, x}(z) - p_{0, x}(z) )
    = \partial_s p_{s, x}(z)$ for some $0 \le s \le 1/\theta$.
    Write $\Gamma(s,x) = \epsilon_r^2 I + s \covq(x)$,
    so that
    \begin{align*}
        p_{s, x}(z)
        &=
        \frac{1}{\left(2 \pi |\Gamma(s,x)|\right)^{d/2}}
        \exp\left(
            -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x))
        \right) ,
    \end{align*}
    and note that if $\lambda_i$ are the eigenvalues of $\covq(x)$
    then $|\Gamma(s, x)| = \prod_i (\epsilon_r^2 + s \lambda_i)$,
    and
    $\partial_s |\Gamma(s, x)| = \sum_i \lambda_i |\Gamma(s, x)| / (\epsilon_r^2 + s \lambda_i)$.
    Therefore,
    \begin{align*}
        \partial_s p_{s, x}(z)
        &=
        \left(
            \meanq(x)^T \Gamma(s,x)^{-1} (z - s \meanq)
            +
            (z - s \meanq)^T
            \Gamma(s,x)^{-1} \covq(x) \Gamma(s,x)^{-1}
            (z - s \meanq)
            -
            \sum_i \frac{\lambda_i}{\epsilon_r^2 + s \lambda_i}
        \right)
        p_{s, x}(z) ,
    \end{align*}
    where $z^T$ is the transpose of $z$.
    Since $\partial_s p_{s, x}(z) / p_{s, x}(z)$ is a polynomial in $z$
    with coefficients that are uniformly bounded for $s \ge 1$,
    it is evident that
    $\sup_z \partial_s p_{\epsilon_r^2, s, x}(z) / p_{\epsilon_\gamma^2}(z) < \infty$
    if and only if
    \begin{align*}
        \frac{1}{2\epsilon_\gamma^2} \|z\|^2
        -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x))
    \end{align*}
    is negative for all $z$ outside a bounded region.
    Since $\inf_z z^T \Gamma(s,x)^{-1} z / \|z\|^2 = 1 / (s \lambda(x) + \epsilon_r^2)$,
    where $\lambda(x) = \sup z^T \covq(x) z / \|z\|^2$ is the largest eigenvalue of $\covq(x)$,
    equation~\eqref{eqn:first_moment_rho} follows from the assumption
    that $\epsilon_r^2 + 2 \sup_x \lambda(x)/\theta < \epsilon_\gamma^2$.
    (Note that we do not yet need the factor of 2.)

    Next we prove equation~\eqref{eqn:second_moment_rho}, in a similar way.
    Again applying Fubini,
    \begin{align*}
        &
        \theta \int_{\mathbb{R}^d}
                \left( \smooth{r} \eta(y) - \smooth{r} \eta(x) \right)^2
        q_\theta(x,dy)
        \\ &\qquad
        =
        \int_{\IR^d} \int_{\IR^d}
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \eta(dv) \eta(dw) ,
    \end{align*}
    and so as before, equation~\eqref{eqn:second_moment_rho} will follow if
    the integrand is bounded by $K p_\gamma(x-w) p_\gamma(x - v)$.
    % \begin{align*}
    %     \theta \int_{\IR^d}
    %         ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
    %         ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
    %     q_\theta(x, dy)
    %     \le K p_\gamma(x-w) p_\gamma(x - v) .
    % \end{align*}
    Now, let $Y_1$, $Y_2$, and $Z$ be independent $d$-dimensional Gaussians with mean zero,
    where $Y_1$ and $Y_2$ have covariance $\sigma^2_r I$,
    and $Z$ has covariance $\covq(x)$.
    Write $p_{s,t,x}(\cdot, \cdot)$ for the joint density of
    $Y_1 + \sqrt{s} Z + s \meanq(x)$ and $Y_2 + \sqrt{t} Z + t \meanq(x)$.
    Then, observe that
    \begin{align*}
        &
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \\ &\qquad
        =
        \theta \left(
            p_{1/\theta, 1/\theta, x}(x-w, x-v)
            - p_{0, 1/\theta, x}(x-w, x-v)
        \right. \\ &\qquad \qquad \left. {}
            - p_{1/\theta, 0, x}(x-w, x-v)
            + p_{0, 0, x}(x-w, x-v)
        \right) 
        \\ &\qquad 
        =
            \partial_s p_{s, 1/\theta, x}(x-w, x-v)
            - \partial_t p_{0, t, x}(x-w, x-v) ,
    \end{align*}
    for some $0 \le s, t \le 1/\theta$.
    As before, $\partial_s p_{s, t, x}(y_1, y_2)$ has polynomial terms multiplying
    $$
        \exp\left(
            - \frac{1}{2} [y_1, y_2[^T \Gamma(s, t, x)^{-1} [y_1, y_2]
        \right) ,
    $$
    where $[y_1, y_2]$ is the $\IR^{2d}$ vector formed by concatenating $y_1$ and $y_2$,
    and $\Gamma(s, t, x)$ is the block matrix
    \begin{align*}
        \Gamma(s, t, x)
        =
        \left[
        \begin{array}{cc}
            \epsilon_r^2 I + s \covq(x) & \sqrt{st} \covq(x) \\
            \sqrt{st} \covq(x) & \epsilon_r^2 I + t \covq(x) \\
        \end{array}
        \right] .
    \end{align*}
    If $\covq(x) u = a u$ for some $a \in \IR$,
    then $[u \sqrt{s}, u \sqrt{t}]$ is an eigenvector of $\Gamma(s, t, x)$
    with eigenvalue $\epsilon^2_r + (s+t) a$,
    and $[u \sqrt{t}, - u \sqrt{s}]$ is an eigenvector of $\Gamma(s, t, x)$
    with eigenvalue 0.
    This implies the largest eigenvalue of $\Gamma(s, t, x)$
    is equal to $\epsilon^2_r + (s+t) \lambda(x)$,
    where $\lambda(x)$ is again the largest eigenvalue of $\covq(x)$.
    Therefore, if $s + t \le 2 / \theta$,
    \begin{align*}
        &
        (\|y_1\|^2 + \|y_2\|^2) / \epsilon^2_\gamma
            - [y_1, y_2[^T \Gamma(s, t, x)^{-1} [y_1, y_2]
        \\ &\qquad \le
        (\|y_1\|^2 + \|y_2\|^2) \left(
            \frac{1}{\epsilon^2_\gamma}
            - \frac{1}{\epsilon^2_r +  2\lambda(x)/\theta}
        \right) ,
    \end{align*}
    which is negative by assumption.
    Therefore, there is a $K$ such that
    \begin{align*}
        \frac{ \left|
            \partial_s p_{s, 1/\theta, x}(x-w, x-v)
            - \partial_t p_{0, t, x}(x-w, x-v)
        \right| }{
            p_{\epsilon^2_\gamma}(x - w)
            p_{\epsilon^2_\gamma}(x - v)
        }
        \le K
    \end{align*}
    for all $\theta > 1$ and all $x$, $v$, and $w \in \IR^d$,
    proving equation~\eqref{eqn:second_moment_rho} and hence the lemma.
\end{proof}

\begin{corollary}
    \label{cor:gamma_bound}
    Under the assumptions of Definition~\ref{def:model_setup},
    including that $\epsilon_{r}^2 + \frac{1}{\theta} < \epsilon_{\gamma}^2$,
    for all $f \in \mathcal{C}^{2}_{0}(\IR^d \times \IR)$
    with bounded first and second derivatives
    and for all $\eta \in \measures$, we have
    \begin{equation}\label{eq: gamma - r radius comparison bound}
    \gamma\big(x,\rho_{\gamma}*\eta(x)\big)
    \left| \theta 
        \int_{\IR^n} 
            \big\{
                f\big(y,\rho_{r}*\eta(y)\big)-f\big(x,\rho_{r}*\eta(x)\big)
            \big\}
        q_{\theta}(x,dy)
    \right|
    \leq C_{\gamma, f}    
    \end{equation}
    for some constant $C_{\gamma,f}$ independent of $\eta$ and $\theta$. 
\end{corollary}

\begin{proof}
Since $f(x_1,...,x_d,w)$ is twice differentiable, we can Taylor expand $f$ around  $(x_1,...,x_d,\rho_{r}*\eta(x))$ up to second order and obtain that for $x=(x_1,..,x_d)$,
\begin{equation}
\begin{aligned}
&~f\big(y,\rho_{r}*\eta(y)\big)-f\big(x,\rho_{r}*\eta(x)\big)\\
=&~ \nabla_x f \big(x,\rho_{r}*\eta(x)\big)
            \cdot (y-x)\\ 
        &~ + \frac{\partial f}{\partial w}
            \big(x,\rho_{r}*\eta(x)\big) \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)\\
        &~ + \frac{1}{2}\sum_{1 \leq i,j \leq d }\frac{\partial^2 f}{\partial x_i x_j}
            \big(z,\rho_{r}*\eta(z)\big) (y_i-x_i)(y_j-x_j)\\
        &~ + \frac{1}{2}\sum_{1 \leq i \leq d}
                \frac{\partial^2 f}{\partial w \partial x_i}
                 \big(z,\rho_{r}*\eta(z)\big) (y_i-x_i)
                    \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)\\
        &~ + \frac{1}{2}\frac{\partial^2 f}{\partial^2 w}
            \big(z,\rho_{r}*\eta(z)\big)
            \left(
            \rho_{r}*\eta(y)-\rho_{r}*\eta(x)
            \right)^2,
\end{aligned}    
\end{equation}
where $z=ty+(1-t)x$ for some $t \in [0,1]$.

Write $||Df||_{\infty}$ and $||D^2f||_{\infty}$
for the suprema of the first and second derivatives of $f$
(which we have assumed are bounded).
Therefore, we have that 
&~\left| 

@@

                f\big(y,\rho_{r}*\eta(y)\big)-f\big(x,\rho_{r}*\eta(x)\big)
                q_{\theta}(x,dy)
\right|\\
\leq &~ ||Df||_{\infty} \left|
                        \theta \int_{\mathbb{R}^d} \sum_{i=1,..,d}(y_i-x_i) q_{\theta}(x,dy)
                        \right|\\ 
        &~ + ||Df||_{\infty} 
            \left| \theta \int_{\mathbb{R}^d}
               \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)
            q_{\theta}(x,dy) \right| \\
        &~ + \frac{1}{2}||D^2f||_{\infty}\left|
                        \theta \int_{\mathbb{R}^d} (y-x)(y-x)^{T} q_{\theta}(x,dy)
                        \right|\\
        &~ + \frac{1}{2}||D^2f||_{\infty}    
                \sum_{1 \leq i \leq d}\left|
                        \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)
                  (y_i-x_i)
                        q_{\theta}(x,dy)
                        \right|

>>>>

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction},
For each value of $N$ and $\theta = \theta(N)$,
we define $(\eta^N_t)_{t \geq 0}$ to be the solution to the Martingale Problem,
with initial condition $\eta^N_0 \in \measures$,
such that for all smooth functions $f \in C^{\infty}_{0}(\mathbb{R}^d)$,
writing $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$
and $q_\theta(x, dy)$ as the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \int\left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
\\ & \qquad \qquad \qquad \qquad {}
        \times \gamma(x, \eta^N_{s}) \eta^N_{s}(dx)
\\ & \qquad \qquad {}
    + \int f(x) F(x, \eta^N_{s}) \eta^N_{s}(dx)
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N_t(f)] =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \langle \gamma(x, \eta^N_{s})
        \int f^2(y)r(y, \eta^N_{s}) q_\theta(x,dy) 
    , \eta^N_{s}(dx) \rangle \\
& \qquad \qquad {}
    + \langle \mu(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}

Note that
since individuals are produced at rate $N \gamma \eta$ but have mass $1/N$ each,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}.
Since we are taking a diffusion limit,
we will have use for the following notation:

@@

The lookdown processes have been carefully constructed so that
if we average over levels in the expression for the lookdown generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we get the generator for the population process
(visible in equation~\eqref{eqn:eta_martingale}
or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (and some boundedness conditions)
in Appendix \ref{sec: Markov Mapping Theorem Application},
the Markov Mapping Theorem \citep[Thm A.2 in][]{etheridge/kurtz:2018}
tells us that for any population process $(\eta^N_t)_{t \ge 0}$
(i.e., solution to the martingale problem of Definition~\ref{defn:mgale_construction})
% of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
there is a lookdown representation $(\lp^N_t)_{t \ge 0}$
(i.e., solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])$
and such that at each $t$, conditioned on $(\eta^N_s){s \le t}$, the levels of $\lp_t^N$
are independently and uniformly distributed on $[0, N]$.
The same property holds in the limit:
for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of Theorem~\ref{nonlocal_convergence}
there is a lookdown process $(\lp^N_t)_{t \ge 0}$
that is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
and such that at each $t$,
$\lp_t$ is conditionally Poisson with Cox measure $\eta_t(dx) \times du$,
conditioned on $(\eta_s)_{s \le t}$.
Furthermore, convergence of the lookdown processes would imply convergence of the population processes
\citep[Thm A.9 in][]{KR}.

@@

We prove tightness of $\eta^N$ as $\mathcal{D}_{\cmeasures}([0,\infty))$-valued processes,
where $\cmeasures$ is the space of finite measures on the one-point compactification of $\IR^d$.
To do this, we first prove that $\sup_{0 \le t \le T} \langle 1, \eta^N_t \rangle$ is tight
(a compact containment condition, Lemma~\ref{lem:eta_compact_containment}).
(The reason we use $\cmeasures$ instead of $\measures$
is so that the set of measures with total mass less than $K$ is compact.)
Next, in Lemma~ we show that the one-dimensional distributions,
$\langle f, \eta^N_t \rangle$ are tight
for each $f \in \mathcal{C}^2_0(\overline{\IR^d})$.
Since this is a dense subset of bounded continuous functions on $\cmeasures$,
together these results allow us to apply Theorem~3.9.1 of \citet{EK}.

@@

To show compact containment, we will need boundedness
of the net per-capita reproduction rate.
This does not follow immediately from our assumption that $F$ is bounded above
because reproduction is not local:
the net per-capita reproduction rate of an individual at $x$ is
$$
\theta \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) - \theta \mu(x, \eta)
=
F(x, \eta) + \gamma(x, \eta) \theta \int_{\IR^d} (r(y, \eta) - r(x, \eta)) q_\theta(x, dy) .
$$
We have assumed that $F$ is bounded,
but we still need to show that our assumptions imply that
the second term is bounded, uniformly as $\theta$ grows.

\begin{lemma}
    \label{lem:gamma_bound}
    Under the assumptions of Definition~\ref{def:model_setup},
    for all $f \in \mathcal{C}^{2}_{0}(\IR^d \times [0,\infty))$,
    there exists a $K_f < \infty$ such that
    for all $\theta > 0$ and all $\eta \in \measures$,
    \begin{equation} \label{eqn:gamma_bound}
    \gamma\big(x,\smooth{\gamma} \eta(x)\big)
    \left| \theta 
        \int_{\IR^n} 
            \left\{
                f(y, \smooth{r} \eta(y)) - f(x, \smooth{r} \eta(x))
            \right\}
        q_{\theta}(x,dy)
    \right|
    \leq C_f     .
    \end{equation}
\end{lemma}

The key assumption here from Definition~\ref{def:model_setup}
will be that the spatial scale over which local density affects birth rate
is larger than the scale over which it affects establishment:
concretely, that $\epsilon_{r}^2 + \frac{2\lambda}{\theta} < \epsilon_{\gamma}^2$,
where $\lambda = \sup_x \sup_{y : \|y\| = 1} y^T \covq(x) y$
is the largest eigenvalue of $\covq(x)$ across all $x$.
In the simple case of $\meanq = 0$ and $\covq = \sigma^2 I$,
the condition is simply that $\epsilon_r^2 + 2\sigma^2 / \theta < \epsilon_\gamma^2$.

\begin{proof}[of Lemma \ref{lem:gamma_bound}]

    Since we have assumed that there is a $K$ for which $m^2 \gamma(x, m) < K$
    for all $x \in \IR^d$ and $m \ge 0$,
    equation~\eqref{eqn:gamma_bound} will follow if we can show that
    \begin{equation*}
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
        \leq
        C_f \left( 1 + \smooth{\gamma} \eta(x)^2 \right) .
    \end{equation*}
    To do this, first note that 
    \begin{align} \label{eqn:q_bound_split}
    &
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(y)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(x, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy) 
        \right| .
    \end{align}

@@

    First we prove \eqref{eqn:first_moment_rho}.
    Recall that $\smooth{r} \eta(x) = \int p_{\epsilon^2_r}(x - w) \eta(dw)$,
    where $p_t$ is the density of a Gaussian with mean 0 and variance $t$,
    so that applying Fubini, \eqref{eqn:first_moment_rho} is
    \begin{align*}
        \left|
        \int_{\IR^d}
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            q_\theta(x, dy)
        \eta(dw)
        \right| .
    \end{align*}
    Write $p_{s, x}(\cdot)$ for the density of a Gaussian
    with mean $s \meanq(x)$ and covariance $\epsilon_r^2 I + s \covq(x)$,
    so that $\int p_{\epsilon_r^2}(y - w) q_\theta(x, dy) = p_{1/\theta, x}(w-x)$.
    It therefore suffices to show that for all $x$ and $w \in \IR^d$,
    there exists $K$ such that
    \begin{align} \label{eqn:goal1}
        \left|
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y-w) - p_{\epsilon_r^2}(x-w) )
            q_\theta(x, dy)
        \right|
        &=
        \theta \left|
            p_{1/\theta, x}(w-x) - p_{0, x}(w-x)
        \right|
        \le
        K p_{\epsilon^2_\gamma}(w-x) .
    \end{align}
    However,
    $\theta( p_{1/\theta, x}(z) - p_{0, x}(z) )
    = \partial_s p_{s, x}(z)$ for some $0 \le s \le 1/\theta$.
    Write $\Gamma(s,x) = \epsilon_r^2 I + s \covq(x)$,
    so that
    \begin{align*}
        p_{s, x}(z)
        &=
        \frac{1}{\left(2 \pi |\Gamma(s,x)|\right)^{d/2}}
        \exp\left(
            -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x))
        \right) ,
    \end{align*}
    and note that if $\lambda_i$ are the eigenvalues of $\covq(x)$
    then $|\Gamma(s, x)| = \prod_i (\epsilon_r^2 + s \lambda_i)$,
    and
    $\partial_s |\Gamma(s, x)| = \sum_i \lambda_i |\Gamma(s, x)| / (\epsilon_r^2 + s \lambda_i)$.
    Therefore,
    \begin{align*}
        \partial_s p_{s, x}(z)
        &=
        \left(
            \meanq(x)^T \Gamma(s,x)^{-1} (z - s \meanq)
            +
            (z - s \meanq(x))^T
            \Gamma(s,x)^{-1} \covq(x) \Gamma(s,x)^{-1}
            (z - s \meanq(x))
        \right. \\ & \qquad \qquad \qquad \left. {}
            -
            \sum_i \frac{\lambda_i}{\epsilon_r^2 + s \lambda_i}
        \right)
        p_{s, x}(z) ,
    \end{align*}
    where $z^T$ is the transpose of $z$.
    This implies that
    \begin{align*}
        \frac{
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y-w) - p_{\epsilon_r^2}(x-w) )
            q_\theta(x, dy)
        }{
            p_{\epsilon^2_\gamma}(w-x) .
        }
        =
        h(z) e^{k(z)},
    \end{align*}
    where $h(z)$ and $k(z)$ are quadratic polynomials in $z$
    whose coefficients depend on $s$ and $x$ but are uniformly bounded for $s \ge 1$
    and $x \in \IR^d$.
    Therefore, \eqref{eqn:goal1} holds for some finite $K$
    if and only if $k(z) < 0$ for all $z$ outside some bounded region.
    This is
    \begin{align*}
        k(z)
        =
        \frac{1}{2\epsilon_\gamma^2} \|z\|^2
        -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x)) ,
    \end{align*}
    which is negative for all $z$ outside a bounded region.
    Since $\inf_z z^T \Gamma(s,x)^{-1} z / \|z\|^2 = 1 / (s \lambda(x) + \epsilon_r^2)$,
    where $\lambda(x) = \sup z^T \covq(x) z / \|z\|^2$ is the largest eigenvalue of $\covq(x)$,
    equation~\eqref{eqn:first_moment_rho} follows from the assumption
    that $\epsilon_r^2 + 2 \sup_x \lambda(x)/\theta < \epsilon_\gamma^2$.
    (Note that we do not yet need the factor of 2.)

    Next we prove equation~\eqref{eqn:second_moment_rho}, in a similar way.
    Again applying Fubini,

@@

                \left( \smooth{r} \eta(y) - \smooth{r} \eta(x) \right)^2
        \\ &\qquad
        =
        \int_{\IR^d} \int_{\IR^d}
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \eta(dv) \eta(dw) ,
    and so as before, equation~\eqref{eqn:second_moment_rho} will follow if
    the integrand is bounded by $K p_\gamma(x-w) p_\gamma(x - v)$.
    % \begin{align*}
    %     \theta \int_{\IR^d}
    %         ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
    %         ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
    %     q_\theta(x, dy)
    %     \le K p_\gamma(x-w) p_\gamma(x - v) .
    % \end{align*}
    Now, let $Y_1$, $Y_2$, and $Z$ be independent $d$-dimensional Gaussians with mean zero,
    where $Y_1$ and $Y_2$ have covariance $\epsilon^2_r I$,
    and $Z$ has covariance $\covq(x)$.
    Write $p_{s,t,x}(\cdot, \cdot)$ for the joint density of
    $Y_1 + \sqrt{s} Z + s \meanq(x)$ and $Y_2 + \sqrt{t} Z + t \meanq(x)$.
    Then, observe that
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \\ &\qquad
        =
        \theta \left(
            p_{1/\theta, 1/\theta, x}(x-w, x-v)
